Classifying Human Driving Behavior via Deep Neural Networks

### 2017-06-16 paper/paper-DrexelFormat.tex
updated paper

% ====================================================================================================================================
% ====================================================================================================================================
\chapter{Background}\label{chap:bg}
% ====================================================================================================================================
% ====================================================================================================================================
This chapter presents the necessary background to understand the experiments presented later in the paper. This chapter is dived into three sections. First, Section \ref{sec:DL} covers basic deep learning concepts including recurrent neural networks and auto-encoders. Second, Section \ref{sec:DLL} compares deep learning libraries: DL4J, Theano, and Tensorflow. Finally, Section \ref{sec:MLDD} briefly discusses existing work in machine learning using for modeling driving data.

@@

Figure \ref{fig:general_AN} shows an general artificial neuron. The threshold is replaced by a {\em bias} term and the {\em subtract} node is replaced by an {\em add} node. In general artificial neurons, the activation function $\phi$ could be an arbitrary function such as step, linear, or sigmoid functions. An general artificial neuron also can be described by the following equation:
$$o = \phi(\vec{x}\cdot\vec{w}+b)$$

@@

In this paper, we used two versions of the collected data set. A first version ({\em raw dataset}) contains 98 of the 100 features collected by the driving simulator (the two features that are removed are time stamps). A second version ({\em filtered dataset}) contains only 23 features. These 23 features were manually selected as are those that are most important for the classification task at hand. Reducing size of features helps save time to train neural network.

\begin{table}[tb]
\centering
\caption{Length of each of the traces used in our experiments.}
\label{tbl:traces}
\begin{tabular}{|l|l|l|l|}
\hline
{\em Trace}   & {\em Driver}    & {\em Track}  & {\em Length} \\ \hline
Trace0  & Expert0   & Track0 & 50029  \\
Trace1  & Expert0   & Track1 & 26375  \\
Trace2  & Expert0   & Track2 & 29629  \\
Trace3  & Expert0   & Track3 & 26298  \\
Trace4  & Expert1   & Track0 & 51295  \\
Trace5  & Expert1   & Track1 & 26674  \\
Trace6  & Expert1   & Track2 & 29680  \\
Trace7  & Expert1   & Track3 & 27075  \\
Trace8  & Inexpert0 & Track0 & 49691  \\
Trace9  & Inexpert0 & Track1 & 30058  \\
Trace10 & Inexpert0 & Track2 & 26441  \\
Trace11 & Inexpert0 & Track3 & 27373  \\
Trace12 & Inexpert1 & Track0 & 47658  \\
Trace13 & Inexpert1 & Track1 & 29380  \\
Trace14 & Inexpert1 & Track2 & 26684  \\
Trace15 & Inexpert1 & Track3 & 27255  \\ \hline
\end{tabular}
\end{table}

@@

\section{LSTM}
The LSTM neural network is used on three different neural network models to classify expert or inexpert drivers because the data has time domain and LSTM gives good performance for serial data. On the paper, LSTM neural networks are built with 16, 32, 64, 128, and 256 hidden neurons. The output from LSTM is sent to an output layer which has two additional neurons. By using softmax, the output from two neurons is classified. If it is [0, 1], it is classified as expert. Otherwise, it is classified as inexpert.

@@

@@

% ####################################################################################################################################
\chapter{Experimental Result}\label{chap:result}
% ####################################################################################################################################
This chapter summarizes results from all experiments. Experiments are divided to two parts: 1) experiments with {\em filtered dataset} 2) experiments with {\em raw dataset}. First part of experiments used {\em first model} which only has LSTM. Each experiments used 5 different number of neurons: 16, 32, 64, 128, and 256. Second part of experiments used {\em second model} which has SAE and {\em third model} which has MAE then compared result of three models.

>>>>

% ====================================================================================================================================
% ====================================================================================================================================
\chapter{Background}\label{chap:bg}
% ====================================================================================================================================
% ====================================================================================================================================
This chapter presents the necessary background to understand the experiments presented later in the paper. This chapter is divided into three sections. First, Section \ref{sec:DL} covers basic deep learning concepts including recurrent neural networks and auto-encoders. Second, Section \ref{sec:DLL} compares deep learning libraries: DL4J, Theano, and Tensorflow. Finally, Section \ref{sec:MLDD} briefly discusses existing work in machine learning using for modeling driving data.

@@

Figure \ref{fig:general_AN} shows a general artificial neuron. The threshold is replaced by a {\em bias} term and the {\em subtract} node is replaced by an {\em add} node. In general artificial neurons, the activation function $\phi$ could be an arbitrary function such as step, linear, or sigmoid functions. A general artificial neuron also can be described by the following equation:
$$o = \phi(\vec{x}\cdot\vec{w}+b)$$

@@

In this paper, we used two versions of the collected data set. A first version ({\em raw dataset}) contains 98 of the 100 features collected by the driving simulator (the two features that are removed are time stamps). A second version ({\em filtered dataset}) contains only 23 features: 'velocity', 'steer', 'brake', 'lane off set', 'throttle', 'stop', and so on. These 23 features were manually selected as are those that are most important for the classification task at hand. The reduced features helps save time to train neural network.

\begin{table}[tb]
\centering
\caption{Sample length on the traces}
\label{tbl:traces}
\begin{tabular}{|l|l|l|l|}
\hline
{\em Trace}   & {\em Driver}    & {\em Track}  & {\em Length} \\ \hline
Trace0  & Expert0   & Track0 & 50029  \\
Trace1  & Expert0   & Track1 & 26375  \\
Trace2  & Expert0   & Track2 & 29629  \\
Trace3  & Expert0   & Track3 & 26298  \\
Trace4  & Expert1   & Track0 & 51295  \\
Trace5  & Expert1   & Track1 & 26674  \\
Trace6  & Expert1   & Track2 & 29680  \\
Trace7  & Expert1   & Track3 & 27075  \\
Trace8  & Inexpert0 & Track0 & 49691  \\
Trace9  & Inexpert0 & Track1 & 30058  \\
Trace10 & Inexpert0 & Track2 & 26441  \\
Trace11 & Inexpert0 & Track3 & 27373  \\
Trace12 & Inexpert1 & Track0 & 47658  \\
Trace13 & Inexpert1 & Track1 & 29380  \\
Trace14 & Inexpert1 & Track2 & 26684  \\
Trace15 & Inexpert1 & Track3 & 27255  \\ \hline
\end{tabular}
\end{table}

@@

\section{LSTM}
The LSTM neural network is used on three different neural network models to classify expert or inexpert drivers because the data has time domain and LSTM gives good performance for serial data. On the paper, LSTM neural networks are built with 16, 32, 64, 128, and 256 hidden neurons. It means that output dimensions of LSTM is 16, 32, 64, 128, and 256 dimensions. The output from LSTM is sent to an output layer which has two additional neurons. By using softmax, the output from two neurons is classified. If it is [0, 1], it is classified as expert. Otherwise, it is classified as inexpert.

@@

\section{Comparison of three models}
Three models are deep neural network which has multiple layers. The {\em first model} has two layers: LSTM and softmax. The {\em second model} has three layers: one encoder, LSTM and softmax. The {\em third model} has five layers: three encoders, LSTM and softmax. Three models contain same structure of LSTM and softmax layer.

The difference between {\em first model} and other models is that the {\em first model} uses {\em filtered dataset} which features are filtered by human but other models reduce features of {\em raw dataset} by AE. The difference between {\em second model} and {\em third model} is that {\em second model} reduces features by SAE but {\em third model} reduces features by MAE. The difference between three models is method for reduction dimensions.

@@

% ####################################################################################################################################
\chapter{Experimental Result}\label{chap:result}
% ####################################################################################################################################
This chapter summarizes results from all experiments. Experiments are divided to two parts: 1) experiments with {\em filtered dataset} 2) experiments with {\em raw dataset}. First part of experiments used {\em first model} which only has LSTM. Each experiments used 5 different number of neurons: 16, 32, 64, 128, and 256. Second part of experiments used {\em second model} which has SAE and {\em third model} which has MAE then compared result of three models.  All experiments were tested four times with four cross validation then the average result are on the chapter.
